Icicle: write once, run once

### 2016-06-09 2016fhpc/section/03-Core.tex
fix typo in S4
Cite anatomy of a loop

The grammar for the Icicle intermediate language is given in Figure~\ref{fig:core:grammar}.
Expressions $PlanX$ include variables, values, applications of primitives and anonymous functions.
% Function definitions and uses are not allowed in expressions here, as their definitions are inlined before converting to query plans.
% Anonymous functions are only allowed as arguments to primitives: they cannot be applied or stored in variables.
The $\mi{Plan}$ itself is split into a five stage \emph{loop anatomy}. First we have the name of the table and the names and element types of each column. The @before@ stage then defines pure values which do not depend on any table data. The @folds@ stage defines element computations and how they are converted to aggregate results. The @after@ stage defines aggregate computations that combine multiple aggregations after the entire table has been processed. Finally, the @return@ stage specifies the output values of the query; a single query will have only one output value, but the result of fusion can have many outputs.

@@

For the @map@ binding, each fold accumulator inside the body of the @group@ construct is associate with its own finite map. The @s@ accumulator is associated with map @gS@, and the @c@ accumulator with @gC@. Each time we receive a row from the table the accumulator associated with the @key@ is updated, using the default value @0@ if an entry for that key is not yet present. After we have processed the entire table each sum with its corresponding count to yield a map of means for each key.
\begin{alltt}
  folds  { gS  : Map Date Real = mapEmpty
           then mapUpdate gS key 0 (\(\lambda\)s. s + value)

>>>>

The grammar for the Icicle intermediate language is given in Figure~\ref{fig:core:grammar}.
Expressions $PlanX$ include variables, values, applications of primitives and anonymous functions.
% Function definitions and uses are not allowed in expressions here, as their definitions are inlined before converting to query plans.
% Anonymous functions are only allowed as arguments to primitives: they cannot be applied or stored in variables.
The $\mi{Plan}$ itself is split into a five stage \emph{loop anatomy}~\cite{Shivers:Anatomy}. First we have the name of the table and the names and element types of each column. The @before@ stage then defines pure values which do not depend on any table data. The @folds@ stage defines element computations and how they are converted to aggregate results. The @after@ stage defines aggregate computations that combine multiple aggregations after the entire table has been processed. Finally, the @return@ stage specifies the output values of the query; a single query will have only one output value, but the result of fusion can have many outputs.

@@

For the @map@ binding, each fold accumulator inside the body of the @group@ construct is associate with its own finite map. The @s@ accumulator is associated with map @gS@, and the @c@ accumulator with @gC@. Each time we receive a row from the table the accumulator associated with the @key@ is updated, using the default value @0@ if an entry for that key is not yet present. After we have processed the entire table we divide each sum with its corresponding count to yield a map of means for each key.
\begin{alltt}
  folds  { gS  : Map Date Real = mapEmpty
           then mapUpdate gS key 0 (\(\lambda\)s. s + value)
